{Gravitational wave sources in our Galactic backyard:}\\{Predictions for BHBH, BHNS and NSNS binaries detectable with LISA

### 2022-03-17 paper/tex/sections/abstract.tex
Updates from Overleaf

For a 4(10)-year LISA mission, we predict between \rangeFourYear{}(\rangeTenYear{}) detections over these variations, out of which 6-154(9-238) are BHBHs, 2-198(3-289) are BHNSs and 3-35(4-57) are NSNSs.
%
We discuss how the variations in the physics assumptions alter the distribution of properties of the detectable systems, even when the detection rates are unchanged. In particular we discuss the observable characteristics such as the chirp mass, eccentricity and sky localisation and how the BHBH, BHNS and NSNS populations can be distinguished, both from each other and from the more numerous double white dwarf population. 
%
We further discuss the possibility of multi-messenger observations of pulsar populations with the Square Kilometre Array (SKA) and assess the benefits of extending the LISA mission.

>>>>

For a 4(10)-year LISA mission, we predict between \rangeFourYear{}(\rangeTenYear{}) detections over these variations, out of which 6-154(9-238) are BHBHs, 2-198(3-289) are BHNSs and 3-35(4-57) are NSNSs.
%
\edit1{We expect that about 50\%(60\%) can be  distinguished from WDWD sources, based on their mass or eccentricity and localisation.  Specifically, for about 10\%(15\%) we expect to be able to determine chirp masses better than 10\%. For 13\%(13\%) we expect sky-localisations better than 1 degree. % 
We discuss the impact of physics variations and prospects for probing the NS-BH mass gap.}  We further discuss the possibility of multi-messenger observations of pulsar populations with the Square Kilometre Array (SKA) and assess the benefits of extending the LISA mission.

### 2022-03-17 paper/tex/sections/introduction.tex
add sentence about GR tests with PSR+BH
Updates from Overleaf

One particularly interesting class of potential LISA sources is BHNSs. With the recent detection of two BHNSs by the LIGO scientific collaboration, the existence of these DCOs has been confirmed \citep{TheLIGOScientificCollaboration+2021}. However, with only two detections (not including the low-confidence candidate GW190426, \citealt{Abbott+2020_GWTC2}, or GW190425, GW190814 and GW190917 which have not been ruled out as BHNSs, \citealt{Abbott+2020_GW190425,Abbott+2020_GW190814, GWTC_2_1}) and no electromagnetic counterparts, the formation rate and properties of BHNSs are still uncertain. Current predictions for the merger rate of BHNSs range across three orders of magnitude \citep[e.g.][]{Abadie+2010, Mandel+2021} so the number of detections in LISA will be important in reducing this uncertainty, thereby refining our understanding of the remnants and evolution of massive stars. Similar to NSNSs, these binaries are also expected to have electromagnetic counterparts. A distinctly exciting possibility is the detection of a pulsar--BH system or millisecond pulsar--BH system \citep{Narayan+1991, Pol+2021}. These systems could be observed not only by LISA, but also radio telescopes such as MeerKAT and the Square Kilometre Array (SKA, \citealt{Dewdney+2009}), which would help to improve the measurement of individual system parameters and to constrain uncertain binary evolution processes \citep[e.g.][]{Pfahl+2005,Chattopadhyay+2020}.

@@

Our paper is structured as follows. In Section~\ref{sec:method}, we describe the methods for synthesising a population of binaries, the variations of physical assumptions that we consider, how we simulate the Milky Way distribution of DCOs and our methods for calculating a detection rate for LISA. We present the main results for our fiducial model in Section~\ref{sec:results}, before exploring the variations in the detectable population when changing our physical assumptions in Section~\ref{sec:variations}. In Section~\ref{sec:discussion} we discuss these results. In Section~\ref{sec:compare_studies}, we compare and contrast our methods and findings to previous work and finish with our conclusions in Section~\ref{sec:conclusion}.

>>>>

One particularly interesting class of potential LISA sources is BHNSs. With the recent detection of two BHNSs by the LIGO scientific collaboration, the existence of these DCOs has been confirmed \citep{TheLIGOScientificCollaboration+2021}. However, with only two detections (not including the low-confidence candidate GW190426, \citealt{Abbott+2020_GWTC2}, or GW190425, GW190814 and GW190917 which have not been ruled out as BHNSs, \citealt{Abbott+2020_GW190425,Abbott+2020_GW190814, GWTC_2_1}) and no electromagnetic counterparts, the formation rate and properties of BHNSs are still uncertain. Similar to NSNSs, these binaries are also expected to have electromagnetic counterparts. A distinctly exciting possibility is the detection of a pulsar--BH system or millisecond pulsar--BH system \citep{Narayan+1991, Pol+2021}. These systems could be observed not only by LISA, but also radio telescopes such as MeerKAT and the Square Kilometre Array (SKA, \citealt{Dewdney+2009}), which would help to improve the measurement of individual system parameters and to constrain uncertain binary evolution processes \citep[e.g.][]{Pfahl+2005,Chattopadhyay+2020}.

@@

Our paper is structured as follows. In Section~\ref{sec:method}, we describe the methods for synthesising a population of binaries, the variations of physical assumptions that we consider, how we simulate the Milky Way distribution of DCOs and our methods for calculating a detection rate for LISA. We present the main results for our fiducial model in Section~\ref{sec:results}, before exploring the variations in the detectable population when changing our physical assumptions in Section~\ref{sec:variations}. \edit1{In Section~\ref{sec:discussion} we discuss these how and whether these sources can be distinguished from the WDWD population and whether the subpopulations of NSNS, BHNS and BHBH can be distinguished from each other. We further discuss prospects for identifying EM counterparts, the role of different formation channels and overall caveats in the modelling approach. }In Section~\ref{sec:compare_studies}, we compare and contrast our methods and findings to previous work and finish with our conclusions in Section~\ref{sec:conclusion}.

### 2022-03-17 paper/tex/sections/methods.tex
Updates from Overleaf

The star formation history of the Milky Way bulge (which we assume here to be dominated by the central bar) has many uncertainties due to the (1) sizeable age measurement uncertainties at large ages in observational studies, (2) complex selection processes affecting the observed age distributions, and (3) formation mechanisms that are still under debate. However, the central bar was shown to contain stars with an extended age range, with most observed stars between $6$ and $12 \unit{Gyr}$ with a younger tail of ages that could come from the subsequent secular growth of the Galactic bar \citep[e.g.,][]{Bovy+2019}. To model the bar's age distribution more realistically than in previous studies (which assume an old bulge coming from a single starburst), we choose to adopt a more extended star formation history using a $\beta(2,3)$ distribution, shifted and scaled such that stars are only formed in the range $[6, 12] \unit{Gyr}$. We show these distributions in the top left panel of Fig.~\ref{fig:galaxy_schematic}.

@@

This scale length accounts for the inside-out growth of the low-\achem~disc and hence is age dependent. We assume $R_d = (1 / 0.43) \unit{kpc}$ for the high-\achem~disc \citep[][Table~1]{Bovy+2016} and $R_d = 1.5 \unit{kpc}$ for the bar component \citep{Bovy+2019}. We show the combination of these distributions in the second panel on the left in Fig.~\ref{fig:galaxy_schematic}.

\textit{Vertical distribution}: Similar to the radial distribution, we use the same single exponential distribution (but with different scale heights) for each component, given by
\begin{equation}\label{eq:galaxy_z}
    p(\abs{z}) = \frac{1}{z_d} \exp\qty(-\frac{z}{z_d}),
\end{equation}
where $z$ is the vertical displacement above the Galactic plane and $z_d$ is the scale height. We set $z_d = 0.3 \unit{kpc}$ for the low-\achem~disc \citep{McMillan+2011} and $z_d = 0.95 \unit{kpc}$ for the high-\achem~disc \citep{Bovy+2016}. For the bar, we set $z_d = 0.2 \unit{kpc}$ \citep{Wegg+15}. We show the combination of these distributions in the bottom left panel of Fig.~\ref{fig:galaxy_schematic}.

>>>>

The star formation history of the bulge/bar of the Milky Way has many uncertainties due to the (1) sizeable age measurement uncertainties at large ages in observational studies, (2) complex selection processes affecting the observed age distributions, and (3) formation mechanisms that are still under debate. However, the central bar (which we assume here to dominate) was shown to contain stars with an extended age range, with most observed stars between $6$ and $12 \unit{Gyr}$ with a younger tail of ages that could come from the subsequent secular growth of the Galactic bar \citep[e.g.,][]{Bovy+2019}. To model the bar/bulge's age distribution more realistically than in previous studies (which assume an old bulge coming from a single starburst), we choose to adopt a more extended star formation history using a $\beta(2,3)$ distribution, shifted and scaled such that stars are only formed in the range $[6, 12] \unit{Gyr}$. We show these distributions in the top left panel of Fig.~\ref{fig:galaxy_schematic}.

@@

This scale length accounts for the inside-out growth of the low-\achem~disc and hence is age dependent. We assume $R_d = (1 / 0.43) \unit{kpc}$ for the high-\achem~disc \citep[][Table~1]{Bovy+2016} and $R_d = 1.5 \unit{kpc}$ for the bulge/bar component \citep{Bovy+2019}. \edit1{Note that in this way we approximate the bulge/bar component as being axi-symmetric, which is sufficient for our purposes.} We show the combination of these distributions in the second panel on the left in Fig.~\ref{fig:galaxy_schematic}.

\textit{Vertical distribution}: Similar to the radial distribution, we use the same single exponential distribution (but with different scale heights) for each component, given by
\begin{equation}\label{eq:galaxy_z}
    p(\abs{z}) = \frac{1}{z_d} \exp\qty(-\frac{z}{z_d}),
\end{equation}
where $z$ is the vertical displacement above the Galactic plane and $z_d$ is the scale height. We set $z_d = 0.3 \unit{kpc}$ for the low-\achem~disc \citep{McMillan+2011} and $z_d = 0.95 \unit{kpc}$ for the high-\achem~disc \citep{Bovy+2016}. For the bulge/bar, we set $z_d = 0.2 \unit{kpc}$ \citep{Wegg+15}. We show the combination of these distributions in the bottom left panel of Fig.~\ref{fig:galaxy_schematic}.
